# Load Libraries
```{r, echo=TRUE, message=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=60)}
setwd("Z:/Users/dnewho/Estes_GeoMx/")
library(RUVSeq)
library(ruv)
library(limma)
library(pheatmap)
library(ggplot2)
library(RColorBrewer)
library(EnhancedVolcano)
library(gridExtra)
library(dplyr)
library(openxlsx)
library(grid)
```

#Get Raw Data & Clean for downstream analyses
```{r}
counts<-read.xlsx("Estes Lab GeoMX DSP CTA experiment with RAW data_03182021.xlsx",
                  sheet = 3, # target count matrix provided by NanoString
                  rowNames = T)

colnames(counts) <- gsub(".|.","_", colnames(counts),fixed=T)
colnames(counts) <- gsub("Geometric.Segment","Geometric_Segment", colnames(counts),fixed=T)

#Get Design
design <-read.xlsx("Estes Lab GeoMX DSP CTA experiment with RAW data_03182021.xlsx",
                  sheet = 1,
                  rowNames = F)
design$SegmentDisplayName <- gsub(" | ","_", design$SegmentDisplayName,fixed=T)
design$SegmentDisplayName <- gsub("Geometric Segment","Geometric_Segment",design$SegmentDisplayName,fixed=T)

row.names(design)=design$SegmentDisplayName

#There are two extra samples in the design that aren't in counts
setdiff(row.names(design),colnames(counts))
#  "35661_CTA_006_Center" "35661_CTA_011_ring2" 

designnew = design[colnames(counts),] # removes the missing two samples
countsnew<-floor(counts) # analysis need integers. Let's round down.

#There are 32 HouseKeeping "Control" genes. Let's use these to help normalize
hk<-c("TFRC","ARMH3","NRDE2","HPRT1","UBB","DNAJC14","ACTB","RPLP0","GPI","POLR2A","EIF2B4","ERCC3","RAB7A","GUSB","OAZ1","SF3A1","MRPL19","TUBB","ABCF1","POLR1B","TPM4","TLK2","PSMB2","HDAC3","PUM1","TMUB2","G6PD","STK11IP","PSMC4","TBC1D10B","TBP")

##  Explore Raw Count distribution
# 1. No Normalization
ses = newSeqExpressionSet(counts = as.matrix(countsnew), phenoData = designnew)
plotRLE(ses, outline=F,xaxt="t", las=3, main='Raw Counts', ylab="RLE")

# 2. Nanostring uses Q3 normalization, i.e. upper quartile. How does this normalization impact count distribution?
ses1<-betweenLaneNormalization(ses,which = "upper")
plotRLE(ses1, outline=F, xaxt="n", main='Q3 Quantile Norm', ylab="RLE")

# 3. What about Full Quantile Normalization?
ses2<-betweenLaneNormalization(ses,which = "full")
plotRLE(ses2, outline=F, xaxt="n", main='Quantile Norm', ylab="RLE")

### Moving into final normalization to incorporate housekeeping genes - Normalization with RUVg. 
# Adapted from :
# Original Blog Posts: https://support.bioconductor.org/p/109778/ & https://support.bioconductor.org/p/123295/ 
# Publication: https://academic.oup.com/bib/article/22/3/bbaa163/5891144 
# Code: https://github.com/bhattacharya-a-bt/CBCS_normalization 


# Raw Counts
normSES = RUVg(ses, 
    hk, 
    k = 2)

#Upper quartile normalization
normSESQ3 = RUVg(ses1, 
    hk, 
    k = 2)

#Full Quantile Normalization
normSESfullQ = RUVg(ses2, 
    hk, 
    k = 2)

plotRLE(normSES, outline=F, xaxt="n", main='Raw RUVg Norm', ylab="RLE")
plotRLE(normSESQ3, outline=F, xaxt="n", main='Q3 RUVg Norm', ylab="RLE")
plotRLE(normSESfullQ, outline=F, xaxt="n", main='Full Quantile RUVg Norm', ylab="RLE")

# Get factors of unwanted variation for future use
W_1 <- normSESfullQ$W_1 
W_2 <- normSESfullQ$W_2
factors <- cbind(W_1,W_2)
```

# Exploring lowly expressed genes. 
# Alternatively, we could start with raw counts and filter based on negative probe expression. This removes genes with consistent low signal across the dataset. Not done here, but worth exploring.
```{r}
# Filter lowly expressed genes based on expression of NegProbe-WTX
y <- DGEList(counts=counts)
A <- rowMeans(y$counts)
A["Negative Probe"] # 154 
isexpr <- A > 154
y <- y[isexpr,]
dim(y)
```

# Standard nCounter uses positive controls, negative controls, and housekeeping genes to normalize. GeoMx seems to use "Q3" normalization.
From nanostring: "Q3 normalization uses the top 25% of expressers to normalize across ROIs/segments, so it is robust to changes in expression of individual genes and ideal for making comparisons across ROIs/segments."
Full quantile normalization looks better here.

```{r}
treatment <- factor(designnew$SegmentLabel, levels = unique(designnew$SegmentLabel))
animalID <- factor(designnew$SlideName, levels = unique(designnew$SlideName))

# Here, we normalize the data using the factors of unwanted variation in the linear model
# We also do not include SIV/anti CD21 treatment status yet, as we want normalization to be blind to the main effect that we will be testing for.
# We then proceed through voom transformation and downstream limma based differential expression testing.
lmdesign <- model.matrix(~0 + treatment + factors)
is.fullrank(lmdesign)
nonEstimable(lmdesign)

y <- DGEList(counts=countsnew)

# Running voom & duplicateCorrelation twice is recommended by limma authors
# https://support.bioconductor.org/p/114663/
eset_voom  <- voom(y, 
                   design=lmdesign,
                   normalize.method = "none",
                   plot = T)


corfit <- duplicateCorrelation(eset_voom,
                              lmdesign,
                              block=animalID) # account for repeated sampling of individuals

eset_voom  <- voom(y, 
                   design=lmdesign,
                   block=animalID,
                   correlation=corfit$consensus.correlation,
                   normalize.method = "none",
                   plot = T)

corfit <- duplicateCorrelation(eset_voom,
                              lmdesign,
                              block=animalID) 

corfit$consensus.correlation

norm_matrix <- eset_voom$E

boxplot(eset_voom$E,ylab = "log2 Expression", main = "Voom Transformation", cex.axis=.6, las=2)

plotMDS(eset_voom, top=300, gene.selection="common")

dat.dist<-dist(t(eset_voom$E))
plot(hclust(dat.dist),labels = designnew$SlideName,cex=.75)

###Densities after normalization
plotDensities((y$counts), legend = FALSE)
plotDensities((eset_voom$E), legend = FALSE)

#write.csv(eset_voom$E,file="EstesLab_GeoMx_NormalizedCounts.csv")
```

```{r}
#### PCA Exploration
contrasting2 = colorRampPalette(rev(c("chocolate2","chocolate1","white","skyblue","deepskyblue3")))(100)


p <- PCAtools::pca(eset_voom$E, 
         metadata = designnew)

PCAtools::biplot(p, x='PC1', y='PC2', 
       lab=NULL, 
       colby='SegmentLabel',
      colkey = c('Center'='darkred','Geometric Segment'='black','ring1'='steelblue','ring2'='mediumpurple'),
#    colkey = c('35661'='grey','35684'='black','37170'='steelblue','37172'='chocolate4'),
     #  colkey = c('FALSE'='steelblue','TRUE'='darkred','FALSE (virions observed)'='black'),

       shape='SlideName',
       pointSize = 4,
       legendPosition = 'right',
       #hline=0, 
       #vline=0,
       axisLabSize = 24,
       legendLabSize = 20,
       legendTitleSize = 24,
       showLoadings = FALSE,
       legendIconSize = 8,
       gridlines.major = FALSE,
       gridlines.minor = FALSE)

ggsave("Estes_PCA_SegmentLabel.pdf", 
       plot=last_plot(), 
       device=cairo_pdf, 
       path=NULL, 
       scale=1, 
       width=12, height=6, 
       dpi=900, limitsize=TRUE)

```

```{r}
# Differential Expression, testing for the effect of SIV and Anti CD21 treatment status across different regions
designnew$Segment_Infection<-paste(designnew$SegmentLabel,designnew$`SIV+`, sep="_") # Grouping variable
designnew$Segment_Infection<-gsub("Geometric Segment","Geometric_Segment",designnew$Segment_Infection)
Segment_Infection <- factor(designnew$treatment, levels = unique(designnew$treatment))

designnew$treatment=paste(designnew$Segment_Infection,designnew$`Treated.(anti-CD21)`,sep="_")


treatment <- factor(designnew$treatment, levels = unique(designnew$treatment))

mm <- model.matrix(~0 + treatment + factors)
is.fullrank(mm)
nonEstimable(mm)


fit <- lmFit(eset_voom,
             block=animalID,
            correlation=corfit$consensus.correlation,
             mm)



# Exploring different contrasts
cont.matrix<-makeContrasts(
     #treatmentGeometric_Segment_NA_CD21_Treatment-treatmentCenter_TRUE_CD21_Treatment,
     #treatmentGeometric_Segment_NA_CD21_Treatment-treatmentring1_TRUE_CD21_Treatment,
     #treatmentGeometric_Segment_NA_CD21_Treatment-treatmentring2_TRUE_CD21_Treatment,
     #treatmentGeometric_Segment_NA_No_Treatment-treatmentCenter_TRUE_No_Treatment,
     #treatmentGeometric_Segment_NA_No_Treatment-treatmentring1_TRUE_No_Treatment,
     #treatmentGeometric_Segment_NA_No_Treatment-treatmentring2_TRUE_No_Treatment,
     #treatmentGeometric_Segment_NA_CD21_Treatment-treatmentCenter_FALSE_CD21_Treatment,
     #treatmentGeometric_Segment_NA_CD21_Treatment-treatmentring1_FALSE_CD21_Treatment,
     #treatmentGeometric_Segment_NA_CD21_Treatment-treatmentring2_FALSE_CD21_Treatment,
     #treatmentGeometric_Segment_NA_No_Treatment-treatmentCenter_FALSE_No_Treatment,
     #treatmentGeometric_Segment_NA_No_Treatment-treatmentring1_FALSE_No_Treatment,
     #treatmentGeometric_Segment_NA_No_Treatment-treatmentring2_FALSE_No_Treatment,
     #treatmentCenter_TRUE_CD21_Treatment-treatmentCenter_TRUE_No_Treatment,
     #treatmentring1_TRUE_CD21_Treatment-treatmentring1_TRUE_No_Treatment,
     #treatmentring2_TRUE_CD21_Treatment-treatmentring2_TRUE_No_Treatment,
     treatmentCenter_TRUE_CD21_Treatment-treatmentCenter_FALSE_CD21_Treatment,
     #treatmentring1_TRUE_CD21_Treatment-treatmentring1_FALSE_CD21_Treatment,
     #treatmentring2_TRUE_CD21_Treatment-treatmentring2_FALSE_CD21_Treatment,
     #treatmentCenter_TRUE_No_Treatment-treatmentCenter_FALSE_No_Treatment,
     #treatmentring1_TRUE_No_Treatment-treatmentring1_FALSE_No_Treatment,
     #treatmentring2_TRUE_No_Treatment-treatmentring2_FALSE_No_Treatment,
     #treatmentCenter_FALSE_CD21_Treatment-treatmentCenter_TRUE_No_Treatment,
     #treatmentring1_FALSE_CD21_Treatment-treatmentring1_TRUE_No_Treatment,
     #treatmentring2_FALSE_CD21_Treatment-treatmentring2_TRUE_No_Treatment,
levels=mm)

cont.matrix<-makeContrasts(
     treatmentGeometric_Segment_NA_CD21_Treatment-treatmentGeometric_Segment_NA_No_Treatment,
 levels=mm)
# 
fit2 <- contrasts.fit(fit,cont.matrix)
fit2 <-eBayes(fit2, robust = TRUE)

results <- decideTests(fit2, 
                       lfc=(0.58), 
                       method="separate", 
                       adjust.method="BH", 
                       p.value=0.05)
summary(results)

dataMatrix <- fit2$coefficients # Extract results of differential expression
sigMask <- dataMatrix * (results**2) # 1 if significant, 0 otherwise
ExpressMatrix <- subset(dataMatrix, rowSums(sigMask) != 0) # filter for significant genes
sigMask <- subset(sigMask, rowSums(sigMask) != 0)
dim(sigMask)
length(sigMask)

png("Estes_globalmodulesheatmap.png",width = 15, height = 12, units = 'in', res = 900)
global_modulesNT <- heatmap.F.rownames(ExpressMatrix, 
                                     cutoff = 3, 
                                     distmethod = "bicor", 
                                     clustermethod = "ward.D", 
                                     clusterdim='row',
                                     colsep =c(3,6,9,12,15,18,21),
                                     cexCol = 2,
                                     key=F,
                                     labRow = FALSE,
                                     #ColSideColors = colcolors,
                                     labCol = c("BCF-Cent","BCF-ring1","BCF-ring2","BCF-Cent","BCF-ring1","BCF-ring2","BCF-Cent","BCF-ring1","BCF-ring2","BCF-Cent","BCF-ring1","BCF-ring2","Cent","ring1","ring2","Cent","ring1","ring2","Cent","ring1","ring2","Cent","ring1","ring2" )
)

dev.off()

for (mod in unique(global_modulesNT$modules)) {
  gm<-as.data.frame(global_modulesNT$modules)
  gm<-gm %>% rownames_to_column('gene')
  tmp<-filter(gm, gm$`global_modulesNT$modules`==mod)
  tmp <- tmp %>% column_to_rownames('gene')
  tmp$`global_modulesNT$modules`<-NULL
  filename<-paste('GlobalModules_EstesGeoMx', mod, 'txt', sep = '.')
  write.table(tmp,file=filename,quote = F)
}
```

# BCFs: Treatment vs No Treatment
```{r}
nm_GS<-norm_matrix[,grep("Geometric", colnames(norm_matrix))]

breaksList=seq(-3, 3, by = .1)
contrasting2 = colorRampPalette(rev(c("deepskyblue3","skyblue","white","chocolate1","chocolate2")))(100)
contrasting3 = colorRampPalette(rev(c('darkblue', 'mediumblue', 'dodgerblue', 'white', 'orange', 'red', 'darkred')))(100)
contrasting4 = colorRampPalette(rev(c("steelblue4","deepskyblue3","skyblue","white","chocolate1","chocolate2","chocolate4")))(100)

BCF_Tx_vs_No<-topTable(fit2, coef=1, adjust.method="BH",number=Inf, sort.by = "p",p.value=0.05,lfc=0.58)
write.csv(BCF_Tx_vs_No,file="BCF_Tx_vs_NoTx.csv")

designGS<-design[colnames(nm_GS),]
annot = data.frame(
  Tx = factor(designGS$`Treated.(anti-CD21)`))
row.names(annot) <- colnames(nm_GS)

ann_colors = list(
    Tx = c('CD21_Treatment' = "black", 'No_Treatment' = "grey"))

png("Estes_BCF_Tx_vsNoTx.png",width = 8, height = 10, units = 'in', res = 900)
pheatmap(nm_GS[row.names(ExpressMatrix),],
         scale="row", 
         color = colorRampPalette(rev(c(name = contrasting3)))(length(breaksList)),
         cluster_rows = T,
         cluster_cols = T,
         show_rownames = F, show_colnames = F,
         border_color="black", 
         annotation_col = annot,
         annotation_colors = ann_colors,
         breaks = breaksList,
         main="BCF: Treatment vs No Tx")
dev.off()
```

# SIV+ vs SIV- Centers
```{r}
cont.matrixSIV_tx<-makeContrasts(
     treatmentCenter_TRUE_CD21_Treatment-treatmentCenter_FALSE_CD21_Treatment,
     #(treatmentCenter_TRUE_CD21_Treatment + treatmentCenter_TRUE_No_Treatment) - (treatmentCenter_FALSE_CD21_Treatment + treatmentCenter_FALSE_No_Treatment),
     #treatmentCenter_TRUE_No_Treatment-treatmentCenter_FALSE_No_Treatment,
levels=mm)

# 
fit2 <- contrasts.fit(fit,cont.matrixSIV_tx)
fit2 <-eBayes(fit2, robust = TRUE)

results <- decideTests(fit2, 
                       lfc=(0.58), 
                       method="separate", 
                       adjust.method="BH", 
                       p.value=0.05)
summary(results)

dataMatrix <- fit2$coefficients # Extract results of differential expression
sigMask <- dataMatrix * (results**2) # 1 if significant, 0 otherwise
ExpressMatrix <- subset(dataMatrix, rowSums(sigMask) != 0) # filter for significant genes

nm_c<-norm_matrix[,grep("Center", colnames(norm_matrix))]
designC<-design[colnames(nm_c),]
annotSIV = data.frame(
  SIV = factor(designC$`SIV+`),
  Tx = factor(designC$`Treated.(anti-CD21)`))
row.names(annotSIV) <- colnames(nm_c)

ann_colorsSIV = list(
    SIV = c('TRUE' = "darkred", 'FALSE' = "thistle3"),
    Tx = c('No_Treatment' = 'grey', 'CD21_Treatment' = 'black'))

png("Estes_TX_SIV.png",width = 8, height = 10, units = 'in', res = 900)
pheatmap(nm_c[row.names(ExpressMatrix),],
         scale="row", 
         color = colorRampPalette(rev(c(name = contrasting3)))(length(breaksList)),
         cluster_rows = T,
         cluster_cols = T,
         show_rownames = F, show_colnames = T,
         border_color="black", 
         annotation_col = annotSIV,
         annotation_colors = ann_colorsSIV,
         breaks = breaksList,
         main="Tx: SIV+ vs SIV-, n=319 genes")
dev.off()
```


# SIV+ Centers vs Rings
```{r}
cont.matrixCentVsRings<-makeContrasts(
     treatmentCenter_TRUE_CD21_Treatment-treatmentring1_TRUE_CD21_Treatment,
      treatmentCenter_TRUE_CD21_Treatment-treatmentring2_TRUE_CD21_Treatment,
     treatmentCenter_TRUE_No_Treatment-treatmentring1_TRUE_No_Treatment,
      treatmentCenter_TRUE_No_Treatment-treatmentring2_TRUE_No_Treatment,
levels=mm)

# 
fit2 <- contrasts.fit(fit,cont.matrixCentVsRings)
fit2 <-eBayes(fit2, robust = TRUE)

results <- decideTests(fit2, 
                       lfc=(0.58), 
                       method="separate", 
                       adjust.method="BH", 
                       p.value=0.05)
summary(results)

dataMatrix <- fit2$coefficients # Extract results of differential expression
sigMask <- dataMatrix * (results**2) # 1 if significant, 0 otherwise
ExpressMatrix <- subset(dataMatrix, rowSums(sigMask) != 0) # filter for significant genes


png("Estes_SIVpos_CentvsRings_EM.png",width = 6, height = 5, units = 'in', res = 900)
pheatmap(ExpressMatrix,
         #scale="row", 
         color = colorRampPalette(rev(c(name = contrasting3)))(length(breaksList)),
         cluster_rows = T,
         cluster_cols = F,
         show_rownames = F, show_colnames = F,
         border_color="black", 
         #annotation_col = annotSIV,
         #annotation_colors = ann_colorsSIV,
         breaks = breaksList,
         main="")
dev.off()
#
nm_c<-norm_matrix[,grep("Center", colnames(norm_matrix))]
nm_r<-norm_matrix[,grep("ring", colnames(norm_matrix))]
nm_cr<-cbind(nm_c,nm_r)

designCR<-design[colnames(nm_cr),]
annotSIV = data.frame(
  SIV = factor(designCR$`SIV+`),
  Tx = factor(designCR$`Treated.(anti-CD21)`),
  Segment = factor(designCR$SegmentLabel))
row.names(annotSIV) <- colnames(nm_cr)

ann_colorsSIV = list(
    SIV = c('TRUE' = "darkred", 'FALSE' = "thistle3"),
    Tx = c('No_Treatment' = 'grey', 'CD21_Treatment' = 'black'),
    Segment = c('Center' = 'cyan', 'ring1' = 'mediumpurple', 'ring2' = 'dodgerblue'))
#

pheatmap(nm_cr[row.names(ExpressMatrix),],
         scale="row", 
         color = colorRampPalette(rev(c(name = contrasting3)))(length(breaksList)),
         cluster_rows = T,
         cluster_cols = T,
         show_rownames = F, show_colnames = T,
         border_color="black", 
         annotation_col = annotSIV,
         annotation_colors = ann_colorsSIV,
         breaks = breaksList,
         main="")
```




# Example:
# Export results from SIV+ vs SIV- centers during Treatment
```{r}
Tx_Center<-topTable(fit, coef=1, adjust.method="BH",number=Inf, sort.by = "p",p.value=0.05,lfc=0.58)
write.csv(Tx_Center,file="Treatment_SIVpos_vs_SIVneg_Center.csv")

```


# Volcano Plot Example
```{r}
BCF_TxvxNo <- topTable(fit2,coef = 1, sort.by='p', adjust.method = "none", p.value=1, number = 1812)

#BCF Tx vs No Tx Volcano
pdf(file="BCF_Tx_vs_NoTx_volcano.pdf",width=6,height=4)
with(BCF_TxvxNo, plot(logFC, -log10(P.Value), pch=21, col="gray90", bg="grey", main="", cex.main=1, 
                 #ylim=c(0,4),
                 #xlim=c(-1.25,1.25), 
                 xlab="Log2FC", ylab="-log10 FDR"))
with(subset(BCF_TxvxNo, logFC >.58 & P.Value <.05), points(logFC, -log10(P.Value), pch=21, col="black", bg="chocolate1"))
with(subset(BCF_TxvxNo, logFC < -.58 & P.Value <.05), points(logFC, -log10(P.Value), pch=21, col="black", bg="steelblue"))
#with(subset(BCF_CenterT, logFC > .45 & P.Value <.003), text(logFC, -log10(P.Value), labels=c("PDCD1","CD6"), font=4,pos=4))
#with(subset(BCF_CenterT, logFC < 0 & P.Value <.001), text(logFC, -log10(P.Value), labels=c("CCL18"), font=4,pos=2))
abline(h=1.3,v=c(-.58,.58),lty=2)
dev.off()
```




# Exporing Nanotrsing Q3 counts
# Nanostring provided normalized counts in the Excel File, output from their analysis suite software.
```{r}
counts<-read.xlsx("Estes Lab GeoMX DSP CTA experiment with RAW data_03182021.xlsx",
                  sheet = 2,
                  rowNames = F)
counts$names<-paste(counts$ScanLabel,counts$ROILabel,counts$SegmentLabel,sep="_")

counts$ScanLabel<-NULL
counts$SegmentLabel<-NULL
counts$ROILabel<-NULL

row.names(counts)=counts$names
counts$names<-NULL
counts=t(counts)

setdiff(row.names(design),colnames(counts))
row.names(designnew) = colnames(counts)
countsnew<-floor(counts) # analysis need integers. Let's round down.

ses = newSeqExpressionSet(counts = as.matrix(countsnew), phenoData = designnew)
plotRLE(ses, outline=F,xaxt="n", main='Nanostring Provided Q3 Counts', ylab="RLE")

```


# High Level DE analysis
```{r}
designnew$SegmentLabel <- gsub("Geometric Segment","Geometric_Segment",designnew$SegmentLabel,fixed=T)

designnew$treatment2<-paste(designnew$SegmentLabel,designnew$`SIV+`,sep="_")
treatment2 <- factor(designnew$treatment2, levels = unique(designnew$treatment2))

mm <- model.matrix(~0 + treatment2 + factors)
is.fullrank(mm)
nonEstimable(mm)


fit <- lmFit(eset_voom,
             mm)

cont.matrixSIV_tx<-makeContrasts(
     treatment2Center_TRUE - treatment2Center_FALSE,
levels=mm)

# 
fit2 <- contrasts.fit(fit,cont.matrixSIV_tx)
fit2 <-eBayes(fit2, robust = TRUE)

results <- decideTests(fit2, 
                       lfc=(0.58), 
                       method="separate", 
                       adjust.method="BH", 
                       p.value=0.05)
summary(results)

dataMatrix <- fit2$coefficients # Extract results of differential expression
sigMask <- dataMatrix * (results**2) # 1 if significant, 0 otherwise
ExpressMatrix <- subset(dataMatrix, rowSums(sigMask) != 0) # filter for significant genes

ExpressMatrix<-as.matrix(ExpressMatrix[c(-15),])

nm_c<-norm_matrix[,grep("Center", colnames(norm_matrix))]
designC<-design[colnames(nm_c),]
annotSIV = data.frame(
  SIV = factor(designC$`SIV+`),
  Tx = factor(designC$`Treated.(anti-CD21)`))
row.names(annotSIV) <- colnames(nm_c)

ann_colorsSIV = list(
    SIV = c('TRUE' = "darkred", 'FALSE' = "thistle3"),
    Tx = c('No_Treatment' = 'grey', 'CD21_Treatment' = 'black'))

png("Estes_AllCenters_SIV.png",width = 8, height = 10, units = 'in', res = 900)
pheatmap(nm_c[row.names(ExpressMatrix),],
         scale="row", 
         color = colorRampPalette(rev(c(name = contrasting3)))(length(breaksList)),
         cluster_rows = T,
         cluster_cols = T,
         show_rownames = T, show_colnames = T,
         border_color="black", 
         annotation_col = annotSIV,
         annotation_colors = ann_colorsSIV,
         breaks = breaksList,
         main="All Centers: SIV+ vs SIV-, n=51 genes")
dev.off()

write.csv(ExpressMatrix,file="AllCenters_SIVposVSSIVneg.csv")
write.fit(fit2, file="AllCenters_SIVpos_vsNeg_limmaresults.csv", digits=3, method="separate", adjust="BH",sep=",")


rnkfile <- data.frame(names = row.names(dataMatrix), dataMatrix)
row.names(rnkfile)<-NULL

library(WebGestaltR)
WebGestaltR(enrichMethod="GSEA",
              organism="hsapiens",
              interestGene=rnkfile,
              projectName = "Centers_SIVpos_vs_neg_BTM",
              saveRawGseaResult=FALSE,
              isOutput = TRUE,
              minNum=5,
              interestGeneType="genesymbol",
              enrichDatabase = c("community-contributed_Hallmark50",
                                 "geneontology_Biological_Process",
                                 "pathway_KEGG"),
              enrichDatabaseType = "genesymbol",
              sigMethod = "fdr", fdrMethod = "BH", fdrThr = 0.05, nThreads=8)


### IL-10 Network HEatmap
il10<-read.table("il10_network_genes.txt")
il10genes<-unique(il10$V1)

png("Estes_AllCenters_SIV_IL10_genes.png",width = 8, height = 10, units = 'in', res = 900)
pheatmap(nm_c[il10genes,],
         scale="row", 
         color = colorRampPalette(rev(c(name = contrasting3)))(length(breaksList)),
         cluster_rows = T,
         cluster_cols = T,
         cutree_rows =  2,
         show_rownames = T, show_colnames = F,
         border_color="black", 
         annotation_col = annotSIV,
         annotation_colors = ann_colorsSIV,
         breaks = breaksList,
         main="All Centers: SIV+ vs SIV-, IL-10 Target genes")
dev.off()
```


# Nanostring Spatial Decon
# Cell Type Deconvolution tool designed by NanoString for spatial omic data.
# Code below was only exploratory and is included for reference.
```{r}
devtools::install_github("Nanostring-Biostats/SpatialDecon",
                         ref = "master", 
                         build_vignettes = FALSE)
library(SpatialDecon)
data("safeTME")
data("safeTME.matches")
signif(safeTME[seq_len(3), seq_len(3)], 2)
heatmap(sweep(safeTME, 1, apply(safeTME, 1, max), "/"),
        labRow = NA, margins = c(10, 5))

# use the NegProbe to estimate per-observation background
per.observation.mean.neg = norm_matrix["Negative Probe", ]
# and define a background matrix in which each column (observation) is the
# appropriate value of per-observation background:
bg = sweep(norm_matrix * 0, 2, per.observation.mean.neg, "+")
dim(bg)


res = spatialdecon(norm = norm_matrix,
                   bg = bg,
                   X = safeTME,
                   align_genes = TRUE)
str(res)
heatmap(res$beta, cexCol = 0.5, cexRow = 0.7, margins = c(10,7))
CellTypeFreq<-res$beta

eset_voomCells  <- voom(CellTypeFreq, 
                   design=lmdesign,
                   normalize.method = "none",
                   plot = T)

fit <- lmFit(eset_voomCells,
             mm)

cont.matrixSIV_tx<-makeContrasts(
     treatment2Center_TRUE - treatment2Center_FALSE,
levels=mm)

fit2 <- contrasts.fit(fit,cont.matrixSIV_tx)
fit2 <-eBayes(fit2, robust = TRUE)

results <- decideTests(fit2, 
                       #lfc=(0.58), 
                       method="separate", 
                       adjust.method="none", 
                       p.value=0.05)
summary(results)

dataMatrix <- fit2$coefficients # Extract results of differential expression
sigMask <- dataMatrix * (results**2) # 1 if significant, 0 otherwise
ExpressMatrix <- subset(dataMatrix, rowSums(sigMask) != 0) # filter for significant genes

```

